---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README/README-fig-"
  )
```


# About

`EconFreedom` contains the economic freedom scores for countries and US states from two sources: 1) the Economic Freedom of the World (EFW) and 2) the Economic Freedom of North America (EFNA) reports.

[Installation](#installation)
<br>
[Datasets](#datasets)

## Installation

```R
# install.package("devtools")
devtools::install_github("acforrester/EconFreedom")
```

## Datasets

### Economic Freedom of the World (EFW)

```{r}
library(tidyverse)

EconFreedom::efwpnl %>% 
  filter(iso_code == "USA")

```

```{r efw_2018, fig.show='asis', echo=FALSE}

map_dat <- EconFreedom::efwpnl %>% 
  filter(year == max(year)) %>% 
  mutate(
    region = case_when(
      iso_code == "GBR" ~ "UK",
      iso_code == "EGY" ~ "Egypt",
      iso_code == "IRN" ~ "Iran",
      iso_code == "KGZ" ~ "Kyrgyzstan",
      iso_code == "COD" ~ "Democratic Republic of the Congo",
      iso_code == "COG" ~ "Republic of Congo",
      iso_code == "KOR" ~ "South Korea",
      iso_code == "RUS" ~ "Russia",
      iso_code == "SVK" ~ "Slovakia",
      iso_code == "SYR" ~ "Syria",
      iso_code == "VEN" ~ "Venezuela",
      iso_code == "USA" ~ "USA",
      iso_code == "YEM" ~ "Yemen",
      TRUE ~ country_name
    )
  )

map_wld <- map_data("world") %>% 
  left_join(., map_dat, by = "region")


ggplot(map_wld, aes(long, lat, group = group))+
  geom_polygon(aes(fill = efw_index ), color = "white")+
  scale_fill_viridis_c(option = "B",
                       name = "EFW Score") +
  theme_classic() + 
  theme(legend.position = "right", 
        axis.text = element_text(size = 8 , colour = "black"),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8, face = "bold")
  ) +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Economic Freedom of the World Score (2018)",
       caption = "Source: Gwartney et al. (2020)"
       )


```



### Economic Freedom of North America (EFNA)

```{r}

library(tidyverse)

EconFreedom::efna %>% 
  filter(usps == "IL")

```

```{r efna_2018, fig.show='asis', echo=FALSE}

map_dat <- EconFreedom::efna %>% 
  filter(year == max(year)) %>%
  mutate(region = tolower(state_name))

map_usa <- map_data("state") %>% 
  left_join(., map_dat, by = c("region"))

ggplot(map_usa, aes(long, lat, group = group))+
  geom_polygon(aes(fill = efna_index), color = "white")+
  scale_fill_viridis_c(option = "B",
                       name = "EFW Score") +
  theme_classic() + 
  theme(legend.position = "right", 
        axis.text = element_text(size = 8 , colour = "black"),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8, face = "bold")
  ) +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Economic Freedom of North America Score (2018)",
       caption = "Source: Stansel et al. (2020)"
       )



```



## Citations

James Gwartney, Robert Lawson, Joshua Hall, and Ryan Murphy (2020). Economic Freedom of the World: 2020 Edition. Fraser Institute.

Dean Stansel, José Torra, and Fred McMahon (2020). Economic Freedom of North America 2020. Fraser Institute.
