---
output: html_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
```

# About

[`Installation`](#installation) | [`Demonstration`](#demonstration) | [`Citations`](#citations)

`EconFreedom` contains two primary datasets on economic freedom:

1. `efwpnl`: The Economic Freedom of the World (EFW) cross-country panel
2. `efna`: The Economic Freedom of the World (EFNA) panel of U.S. states

## Contents

## Installation

```{r Init}
# install.package("devtools")
devtools::install_github("acforrester/EconFreedom")

```

## Demonstration

```{r Load_Pkgs}
# load packages to use
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,plotly,DT)

```

### Economic Freedom of the World (EFW)

```{r EFW_Datatable}
# load the EFW panel
EconFreedom::efwpnl %>% 
  # keep
  arrange(-year, iso_code) %>%
  # print data
  DT::datatable(., 
                rownames = FALSE,
                extensions = c("Buttons"), 
                options = list(
                  dom = "Bfrtip",
                  buttons = c("csv", "excel", "pdf")
                  )
                )

```

```{r EFW_Map, fig.show='asis', echo=FALSE}
# a bit of cleaning for the shapefile
map_dat <- EconFreedom::efwpnl %>% 
  filter(year == max(year)) %>% 
  mutate(
    region = case_when(
      iso_code == "GBR" ~ "UK",
      iso_code == "EGY" ~ "Egypt",
      iso_code == "IRN" ~ "Iran",
      iso_code == "KGZ" ~ "Kyrgyzstan",
      iso_code == "COD" ~ "Democratic Republic of the Congo",
      iso_code == "COG" ~ "Republic of Congo",
      iso_code == "KOR" ~ "South Korea",
      iso_code == "RUS" ~ "Russia",
      iso_code == "SVK" ~ "Slovakia",
      iso_code == "SYR" ~ "Syria",
      iso_code == "VEN" ~ "Venezuela",
      iso_code == "USA" ~ "USA",
      iso_code == "YEM" ~ "Yemen",
      TRUE ~ country_name
    )
  )

# load a WGS shapefile
map_wld <- map_data("world") %>% 
  # join geodata
  left_join(., map_dat, by = "region") %>% 
  # round EFW
  mutate(EFW = round(efw_index,2))

# plotting
p <- ggplot(map_wld, aes(long, lat, 
                         group = group, 
                         text = paste0(region, " (", iso_code, ")")))+
  geom_polygon(aes(fill = EFW ), color = "white")+
  scale_fill_gradient(low = "#fc0303",
                      high = "#0b03fc",
                      na.value = "white")+
  theme_classic() + 
  theme(legend.position = "right", 
        axis.text = element_text(size = 8 , colour = "black"),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8, face = "bold")
  ) +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Economic Freedom of the World Score (2019)",
       caption = "Source: Gwartney et al. (2021)"
       )

# now call `plotly`
ggplotly(p, tooltip = c("text", "fill"))

```



### Economic Freedom of North America (EFNA)

```{r EFNA_Datatable}
# load the EFNA panel
EconFreedom::efna %>% 
  # keep
  arrange(-year, stfips) %>%  
  # print data
  DT::datatable(., 
                rownames = FALSE,
                extensions = c("Buttons"), 
                options = list(
                  dom = "Bfrtip",
                  buttons = c("csv", "excel", "pdf")
                  )
                )

```

```{r EFNA_map, fig.show='asis', echo=FALSE}
# prep EFNA for a map
map_dat <- EconFreedom::efna %>% 
  filter(year == max(year)) %>%
  mutate(region = tolower(state_name),
         EFNA   = round(efna_index, 2))

# get a shapefile and join EFNA
map_usa <- map_data("state") %>% 
  left_join(., map_dat, by = c("region"))

# plotting
p <- ggplot(map_usa, aes(long, lat, group = group))+
  geom_polygon(aes(fill = EFNA), color = "white")+
  scale_fill_viridis_c(option = "B",
                       name = "EFW Score") +
  theme_classic() + 
  theme(legend.position = "right", 
        axis.text = element_text(size = 8 , colour = "black"),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8, face = "bold")
  ) +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Economic Freedom of North America Score (2018)",
       caption = "Source: Stansel et al. (2020)"
       )

# now call `plotly`
ggplotly(p, tooltip = c("text", "fill"))


```

## Citations

James Gwartney, RobertLawson, Joshua Hall, and Ryan Murphy (2021). *Economic Freedom Dataset, published in Economic Freedom of the World: 2021 Annual Report*. Fraser Institute. [www.fraserinstitute.org/economic-freedom/dataset](www.fraserinstitute.org/economic-freedom/dataset)

Dean Stansel, José Torra, and Fred McMahon (2020). *Economic Freedom of North America 2020*. Fraser Institute. [https://www.fraserinstitute.org/studies/economic-freedom-of-north-america-2020](https://www.fraserinstitute.org/studies/economic-freedom-of-north-america-2020)
